---
output: github_document
---

# Manual for the use of the compIntegrate package

## Publication

The underlying method of the RCM package is described in detail in the following article: ["A unified framework for unconstrained and constrained ordination of microbiome read count data"](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0205474).

## Install and load packages

This repo contains R-code to fit and plot the RC(M)-models augmented with the negative binomial. The basic usage is demonstrated here, for more advanced examples see the [RCM manual](http://users.ugent.be/~shawinke/RCMmanual) and the help pages of _RCM()_, *RCM_NB()*, and _plot.RCM()_ functions.

The package can be installed from BioConductor and loaded using the following commands:

```{r installRCMpackageBioConductor, eval = FALSE}
library(BiocManager)
install("compIntegrate", update = FALSE)
```

```{r loadRCMpackage}
suppressPackageStartupMessages(library(compIntegrate))
cat("compIntegrate package version", as.character(packageVersion("compIntegrate")), "\n")
```

Alternatively, the latest version can be installed directly from this GitHub repo as follows:

```{r installRCMpackage, eval = FALSE}
library(devtools)
install_github("CenterForStatistics-UGent/RCM")
```
