---
output: github_document
---

# Manual for the use of the combi package

## Install and load packages

This repo contains R-code to fit and plot the mode-based integration models for compositional omics data using the _combi_ package (Compositional Omics Model-Based Integration). The basic usage is demonstrated here.

The package can be installed from CRAN and loaded using the following commands:

```{r installcombipackageCRAN, eval = FALSE}
install.packages("combi")
```

```{r loadcombipackage}
suppressPackageStartupMessages(library(combi))
cat("combi package version", as.character(packageVersion("combi")), "\n")
```

Alternatively, the latest version can be installed directly from this GitHub repo as follows:

```{r installcombipackage, eval = FALSE}
library(devtools)
install_github("CenterForStatistics-UGent/combi")
```

## Unconstrained integration

For an unconstrained ordination, a named list of data matrices with overlapping samples must be supplied. In addition, information on the required distribution ("quasi" for quasi-likelihood fitting, "gaussian" for normal data) and compositional nature should be supplied.

```{r unconstr}
data(Zhang)
microMetaboInt = combi(
 list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
 distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
 logTransformGaussian = FALSE)
```

A simple plot function is available for the result, for samples and shapes, a data frame should also be supplied

```{r simplePlot}
plot(microMetaboInt)
```

```{r colourPlot}
plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX")
```

## Constrained integration

For a constrained ordination also a data frame of sample variables should be supplied

```{r constr}
microMetaboIntConstr = combi(
     list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
     distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
     logTransformGaussian = FALSE, covariates = zhangMetavars)
```

```{r colourPlotConstr}
plot(microMetaboIntConstr, samDf = zhangMetavars, samCol = "ABX")
```

## Diagnostics

Convergence of the iterative algorithm can be assessed as follows:

```{r convPlot}
convPlot(microMetaboInt)
```

Influence of the different views can be investigated through

```{r inflPlot}
inflPlot(microMetaboInt, samples = 1:20, plotType = "boxplot")
```
