---
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Manual for the combi pacakage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# combi package: vignette

\setcounter{tocdepth}{5}
\tableofcontents

# Introduction

This package implements a novel data integration model for sample-wise integration of different views. It accounts for compositionality and employs a non-parametric mean-variance trend for sequence count data. The resulting model can be conveniently plotted to allow for explorative visualization of variability shared over different views.

<!-- # Publication -->

<!-- The underlying method of the combi package is described in detail in the following article:  -->

# Installation

The package can be installed and loaded using the following commands:

```{r load-packages, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache = FALSE, autodep = TRUE, warning = FALSE, 
                      message = FALSE, echo = TRUE, eval = TRUE, 
                      tidy = TRUE, fig.width = 9, fig.height = 6, purl = TRUE, 
                      fig.show = "hold", cache.lazy = FALSE)
palStore = palette()
#Load all fits, to avoid refitting every time rebuilding the vignette
load(system.file("extdata", "zhangFits.RData", package = "combi"))
```

```{r install, eval = FALSE}
library(BiocManager)
BiocManager::install("combi", update = FALSE)
```

```{r installDevtools, eval = FALSE}
library(devtools)
install_github("CenterForStatistics-UGent/combi")
```

```{r loadcombipackage}
suppressPackageStartupMessages(library(combi))
cat("combi package version", 
    as.character(packageVersion("combi")), "\n")
```

```{r loadData}
data(Zhang)
```

## Unconstrained integration

For an unconstrained ordination, a named list of datasets with overlapping samples must be supplied. The datasets can currently be supplied as a raw data matrix (with features in the columns), or as a phyloseq, SummarizedExperiment or ExpressionSet object. In addition, information on the required distribution ("quasi" for quasi-likelihood fitting, "gaussian" for normal data) and compositional nature (TRUE/FALSE) should be supplied

```{r unconstr}
microMetaboInt = combi(
 list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
 distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
 logTransformGaussian = FALSE)
```

One can print basic infor about the ordination

```{r show}
microMetaboInt
```

A simple plot function is available for the result, for samples and shapes, a data frame should also be supplied

```{r simplePlot}
plot(microMetaboInt)
```

```{r colourPlot}
plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX")
```

By default, only the most important features (furthest away from the origin) are shown. To show all features, one can resort to point cloud plots or density plots as follows:

```{r cloudPlot}
plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX", 
     featurePlot = "points")
```

```{r denPlot}
plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX", 
     featurePlot = "density")
```

The drawback is that now no feature labels are shown.

## Adding projections

As an aid to interpretation of compositional views, links between features can be plotted and projected onto samples by providing their names or approximate coordinates

```{r projections}
#First define the plot, and return the coordinates
mmPlot = plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX", returnCoords = TRUE, featNum = 10)
#Providing feature names, and sample coordinates, but any combination is allowed
addLink(mmPlot, links = cbind("Staphylococcus_819c11","OTU929ffc"), Views = 1, samples = c(0,1))
```

## Coordinates

Finally, one can extract the coordinates for use in third-party software

```{r extractCoords}
coords = extractCoords(microMetaboInt, Dim = c(1,2))
```

## Constrained integration

For a constrained ordination also a data frame of sample variables should be supplied

```{r constr}
microMetaboIntConstr = combi(
     list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
     distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
     logTransformGaussian = FALSE, covariates = zhangMetavars)
```

Also here we can get a quick overview

```{r printConstr}
microMetaboIntConstr
```

and plot the ordination

```{r colourPlotConstr}
plot(microMetaboIntConstr, samDf = zhangMetavars, samCol = "ABX")
```

## Diagnostics

Convergence of the iterative algorithm can be assessed as follows:

```{r convPlot}
convPlot(microMetaboInt)
```

Influence of the different views can be investigated through

```{r inflPlot}
inflPlot(microMetaboInt, samples = 1:20, plotType = "boxplot")
```

# FAQ

## Why are not all my samples shown in the constrained ordination?

Confusion often arises as to why less distinct sample dots are shown than there are samples in the constrained ordination. This occurs when few, categorical constraining variables are supplied as below.

```{r linFewVars, eval = FALSE}
#Linear with only 2 variables
microMetaboIntConstr2Vars = combi(
     list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
     distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
     logTransformGaussian = FALSE, covariates = zhangMetavars[, c("Sex", "ABX")])
```

Every constrained sample score is a linear combination of constraining variables, and in this case there are only 2x2=4 distinct sample scores possible, leading to the sample dots from the same gender and treatment to be plotted on top of each other.

```{r plotFewVars}
plot(microMetaboIntConstr2Vars, samDf = zhangMetavars, samCol = "ABX")
```

In general, it is best to include all measured sample variables in a constrained analysis, and let the combi-algorithm find out which ones are the most important drivers of variability.

# Session info

This vignette was generated with following version of R:

```{r sessionInfo}
sessionInfo()
```
