---
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Manual for the compIntegrate pacakage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# compIntegrate package: vignette

\setcounter{tocdepth}{5}
\tableofcontents

# Introduction

This package implements a novel data integration model for sample-wise integration of different views. It accounts for compositionality and employs a non-parametric mean-variance trend for sequence count data. The resulting model can be conveniently plotted to allow for explorative visualization of variability shared over different views.

<!-- # Publication -->

<!-- The underlying method of the compIntegrate package is described in detail in the following article:  -->

# Installation

The package can be installed and loaded using the following commands:

```{r load-packages, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(cache = FALSE, autodep = TRUE, warning = FALSE, 
                      message = FALSE, echo = TRUE, eval = TRUE, 
                      tidy = TRUE, fig.width = 9, fig.height = 6, purl = TRUE, 
                      fig.show = "hold", cache.lazy = FALSE)
palStore = palette()
#Load all fits, to avoid refitting every time rebuilding the vignette
#load(system.file("fits", "zellerFits.RData", package = "RCM"))
```

```{r install, eval = FALSE, echo = FALSE}
library(BiocManager)
install("compIntegrate", update = FALSE)
```

```{r installDevtools, eval = FALSE}
library(devtools)
install_github("CenterForStatistics-UGent/compIntegrate")
```

```{r loadcompIntpackage}
suppressPackageStartupMessages(library(compIntegrate))
cat("compIntegrate package version", 
    as.character(packageVersion("compIntegrate")), "\n")
```

## Unconstrained integration

For an unconstrained ordination, a named list of data matrices with overlapping samples must be supplied. In addition, information on the required distribution ("quasi" for quasi-likelihood fitting, "gaussian" for normal data) and compositional nature (TRUE/FALSE) should be supplied

```{r unconstr}
data(Zhang)
microMetaboInt = compInt(
 list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
 distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
 logTransformGaussian = FALSE)
```

A simple plot function is available for the result, for samples and shapes, a data frame should also be supplied

```{r simplePlot}
plot(microMetaboInt)
```

```{r colourPlot}
plot(microMetaboInt, samDf = zhangMetavars, samCol = "ABX")
```

## Constrained integration

For a constrained ordination also a data frame of sample variables should be supplied

```{r constr}
microMetaboIntConstr = compInt(
     list("microbiome" = zhangMicrobio, "metabolomics" = zhangMetabo),
     distributions = c("quasi", "gaussian"), compositional = c(TRUE, FALSE),
     logTransformGaussian = FALSE, covariates = zhangMetavars)
```

```{r colourPlotConstr}
plot(microMetaboIntConstr, samDf = zhangMetavars, samCol = "ABX")
```

## Diagnostics

Convergence of the iterative algorithm can be assessed as follows:

```{r convPlot}
convPlot(microMetaboInt)
```

Influence of the different views can be investigated through

```{r inflPlot}
inflPlot(microMetaboInt, samples = 1:20, plotType = "boxplot")
```

# Session info

This vignette was generated with following version of R:

```{r sessionInfo}
sessionInfo()
```
